Distributed Decision Making for Dynamic Formation of Web Services Communities

### 2015-01-23 papers/webservice-coalition-learning-omar-template/elsarticle-template-num.tex
changes

To effectively train web services, we used real data set to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. The data set has been extracted by real-world QoS evaluation results from 142 users on 4,532 Web services on 64 different time slots. Combining these data based on each web service point of view on different time slots, we acquired 5 different unique features for 4,532 web services. By engineering and extracting these features, we gathered functional and cooperative features for both individual web services and communities in different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in the training model. Using cross validation sets, web services are able to compute the optimal hypothesis function (using logistic regression) that can be used to predict outcomes of cooperative work with other individual web services or communities. Our findings show that web services equipped with DDM have by far better outcomes in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

@@

The goal of all parties involved in the community is to maximize their long-term outcome while they are operating as part of the community. Web services need to be equipped with a selection strategy to choose from all of the different possible collaboration groups they can form as well as an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment with different possible collaborative groups and other web services in order to estimate their gain over time. However, with a high number of possible communities and many other business restrictions, it is not possible to test collaboration with random web services. Even if a linear approximate function for estimating utility based on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of the possible number of partitions of web services into communities would make any brute-force type algorithm for the best community selection strategy intractable and impractical in real-world application settings.

@@

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of web services, about which the web service has recently gathered information. The less value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of web services, about which the web service has recently gathered information. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

@@

In this section, we describe our data extraction process and the methodology used to equip web services with a decision making mechanism. In this methodology, we first present the data extraction and engineering process and then we evaluate the decision making mechanism for web services in community settings.

@@

\subsubsection{Web Service data}\label{sss:webservices}

We used a data set extracted from running web services. In this data set, each web service is associated with a number of features that reflect its functionality. These web services operate in an online environment and are continuously assigned tasks to handle. Sometimes, they pass tasks around, distributing sub-tasks among themselves and handling tasks in a collaborative effort. To engage in communities, we add additional features that reflect web services' utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for a community to be formed. Since we have 64 different time slots of extracted features for each web service, we can build a synthetic data set that contains features of a large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for other web services. To train our model, we build a decision tree as a benchmark for our decision making mechanism. The decision tree is built by training the real data obtained from operating web services and extracting features related to their performance, either alone or as part of a joint effort with other web services. As a result, we are able to classify web services as either single or joined and propose the communities that are best fitted to the web services' requirements.

>>>>

To effectively generate the decision model for web services, we used real data set to extract web services' individual characteristics and used them to measure outcomes when they cooperate with one another. The data set has been extracted by real-world QoS evaluation results from 142 users on 4,532 Web services on 64 different time slots. Combining these data based on each web service point of view on different time slots, we acquired 5 different unique features for 4,532 web services. By engineering and extracting these features, we gathered functional and cooperative features for both individual web services and communities in different time slots. We were able to investigate the path a web service might take to achieve the best utility out of effective interaction with others. All paths and the outcomes are labeled to be utilized in the training model. Using cross validation sets, web services are able to compute the optimal hypothesis function (using logistic regression) that can be used to predict outcomes of cooperative work with other individual web services or communities. Our findings show that web services equipped with DDM have by far better outcomes in the form of cooperative service providers than the ones that either do not cooperate or randomly find communities to join.

@@

The goal of all parties involved in the community is to maximize their long-term outcome while they are operating as part of the community. Web services need to be equipped with a selection strategy to choose from all of the different possible collaboration groups they can form as well as an estimation method for evaluating the long-term gain of joining different possible communities. Web services need to experiment with different possible collaborative groups and other web services in order to estimate their gain over time. However, with a high number of possible communities, it is not possible to test collaboration with random web services. Even if a linear approximate function for estimating utility based on community web services' parameters is adopted, the exponential \footnote{Bell number: $http://en.wikipedia.org/wiki/Bell_number$} growth rate of the possible number of partitions of web services into communities would make any brute-force type algorithm for the best community selection strategy intractable and impractical in real-world application settings.

@@

\begin{itemize}
  \item \emph{External Parameter 1} $(exp_1)$. is an estimate of how close the web service's \emph{execution time} is to other web services. It is the difference between a web service's \emph{execution time} metric and the maximum value of execution time of all the other web services, about which the web service has information of them. The less value means web service has better execution time compared to other web services.
	\begin{equation}\label{exp_1:f}
		exp_1 = Et_{i} - Et_{max}
	\end{equation}
	\item \emph{External Parameter 2} $(exp_2)$ is a comparison of the web service's rate of performing tasks to other competing web services. It is the difference between a web service's \emph{throughput} metric and the maximum value of throughput of all the other web services,, about which the web service has information of them. The less value means web service has better throughput compared to other web services.
	\begin{equation}\label{exp_2Lf}
		exp_2 = Th_{max} - Th_{i}
	\end{equation}
\end{itemize}

@@

In this section, we describe our data extraction process and the methodology used to equip web services with a decision making mechanism. In this methodology, we first present the data extraction and engineering process and then we evaluate the decision making mechanism for web services in community settings. Figure \ref{fig_steps} summarizes the steps performed in DDM from the input data to generation of decision making profiles for web services.

\begin{figure}%[!t]
\centerline{\includegraphics[width=1.75in]{figures/steps.eps}}
\caption{A summary of DDM decision profile generation steps}
\label{fig_steps}
\end{figure}

@@

\subsubsection{Input Web Services Data}\label{sss:webservices}

%We used a data set extracted from running web services. In this data set, 
Each web service is associated with a number of features that reflect its functionality. These web services operate in an online environment and are continuously assigned tasks to handle. Sometimes, they pass tasks around, distributing sub-tasks among themselves and handling tasks in a collaborative effort. To engage in communities, we add additional features that reflect web services' utility as a result of joining other web services to form a community. The additional features are computed based on two assumptions that we adopt for a community to be formed. Our raw data set provides us with 64 different time slots of extracted features for each web service. Using this data we can build a synthetic data set that contains features of a large number of web services in a sequence of time. Therefore, we train a learning model that adopts the trend of joining a community and use the model to predict/find the appropriate community for other web services. To train our model, we build a decision tree as a benchmark for our decision making mechanism. The decision tree is built by training the real data obtained from operating web services and extracting features related to their performance, either alone or as part of a joint effort with other web services. As a result, we are able to classify web services as either single or joined and propose the communities that are best fitted to the web services' requirements.
